---
title: "STAT 4620 Project Data Preparation"
author: "Luke Dinan"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Load Libraries/Data

```{r}
library(tidyverse)
library(magrittr)
```

```{r}
load("Ames/ames.Rdata")
```

##### Replace NA Categorical

Training Data: replace categorical NA with more descriptive value
```{r}
train$Alley <- ifelse(is.na(train$Alley), "None", train$Alley)
train$Fireplace.Qu <- ifelse(is.na(train$Fireplace.Qu), "None", train$Fireplace.Qu)
train$Bsmt.Qual <- ifelse(is.na(train$Bsmt.Qual), "No Basement", train$Bsmt.Qual)
train$Bsmt.Cond <- ifelse(is.na(train$Bsmt.Cond), "No Basement", train$Bsmt.Cond)
train$Bsmt.Exposure <- ifelse(is.na(train$Bsmt.Exposure), "No Basement", train$Bsmt.Exposure)
train$BsmtFin.Type.1 <- ifelse(is.na(train$BsmtFin.Type.1), "No Basement", train$BsmtFin.Type.1)
train$BsmtFin.Type.2 <- ifelse(is.na(train$BsmtFin.Type.2), "No Basement", train$BsmtFin.Type.2)
train$Pool.QC <- ifelse(is.na(train$Pool.QC), "No Pool", train$Pool.QC)
train$Garage.Type <- ifelse(is.na(train$Garage.Type), "No Garage", train$Garage.Type)
train$Garage.Finish <- ifelse(is.na(train$Garage.Finish), "No Garage", train$Garage.Finish)
train$Garage.Qual <- ifelse(is.na(train$Garage.Qual), "No Garage", train$Garage.Qual)
train$Garage.Cond <- ifelse(is.na(train$Garage.Cond), "No Garage", train$Garage.Cond)
train$Fence <- ifelse(is.na(train$Fence), "No Fence", train$Fence)
train$Misc.Feature <- ifelse(is.na(train$Misc.Feature), "None", train$Misc.Feature)
```

Testing Data: replace categorical NA with more descriptive value
```{r}
test$Alley <- ifelse(is.na(test$Alley), "None", test$Alley)
test$Fireplace.Qu <- ifelse(is.na(test$Fireplace.Qu), "None", test$Fireplace.Qu)
test$Bsmt.Qual <- ifelse(is.na(test$Bsmt.Qual), "No Basement", test$Bsmt.Qual)
test$Bsmt.Cond <- ifelse(is.na(test$Bsmt.Cond), "No Basement", test$Bsmt.Cond)
test$Bsmt.Exposure <- ifelse(is.na(test$Bsmt.Exposure), "No Basement", test$Bsmt.Exposure)
test$BsmtFin.Type.1 <- ifelse(is.na(test$BsmtFin.Type.1), "No Basement", test$BsmtFin.Type.1)
test$BsmtFin.Type.2 <- ifelse(is.na(test$BsmtFin.Type.2), "No Basement", test$BsmtFin.Type.2)
test$Pool.QC <- ifelse(is.na(test$Pool.QC), "No Pool", test$Pool.QC)
test$Garage.Type <- ifelse(is.na(test$Garage.Type), "No Garage", test$Garage.Type)
test$Garage.Finish <- ifelse(is.na(test$Garage.Finish), "No Garage", test$Garage.Finish)
test$Garage.Qual <- ifelse(is.na(test$Garage.Qual), "No Garage", test$Garage.Qual)
test$Garage.Cond <- ifelse(is.na(test$Garage.Cond), "No Garage", test$Garage.Cond)
test$Fence <- ifelse(is.na(test$Fence), "No Fence", test$Fence)
test$Misc.Feature <- ifelse(is.na(test$Misc.Feature), "None", test$Misc.Feature)
```


##### Impute Continuous Nulls

Training Data: impute missing values with the median
```{r}
train$Lot.Frontage[is.na(train$Lot.Frontage)]<-median(train$Lot.Frontage,na.rm=TRUE)
train$Garage.Yr.Blt[is.na(train$Garage.Yr.Blt)]<-median(train$Garage.Yr.Blt,na.rm=TRUE)
train$Mas.Vnr.Area[is.na(train$Mas.Vnr.Area)]<-median(train$Mas.Vnr.Area,na.rm=TRUE)
train$Bsmt.Full.Bath[is.na(train$Bsmt.Full.Bath)]<-median(train$Bsmt.Full.Bath,na.rm=TRUE)
train$Bsmt.Half.Bath[is.na(train$Bsmt.Half.Bath)]<-median(train$Bsmt.Half.Bath,na.rm=TRUE)
train$BsmtFin.SF.1[is.na(train$BsmtFin.SF.1)]<-median(train$BsmtFin.SF.1,na.rm=TRUE)
train$BsmtFin.SF.2[is.na(train$BsmtFin.SF.2)]<-median(train$BsmtFin.SF.2,na.rm=TRUE)
train$Bsmt.Unf.SF[is.na(train$Bsmt.Unf.SF)]<-median(train$Bsmt.Unf.SF,na.rm=TRUE)
train$Total.Bsmt.SF[is.na(train$Total.Bsmt.SF)]<-median(train$Total.Bsmt.SF,na.rm=TRUE)
train$Garage.Cars[is.na(train$Garage.Cars)]<-median(train$Garage.Cars,na.rm=TRUE)
train$Garage.Area[is.na(train$Garage.Area)]<-median(train$Garage.Area,na.rm=TRUE)
```

Test Data: impute missing values with the median
```{r}
test$Lot.Frontage[is.na(test$Lot.Frontage)]<-median(test$Lot.Frontage,na.rm=TRUE)
test$Garage.Yr.Blt[is.na(test$Garage.Yr.Blt)]<-median(test$Garage.Yr.Blt,na.rm=TRUE)
test$Mas.Vnr.Area[is.na(test$Mas.Vnr.Area)]<-median(test$Mas.Vnr.Area,na.rm=TRUE)
test$Bsmt.Full.Bath[is.na(test$Bsmt.Full.Bath)]<-median(test$Bsmt.Full.Bath,na.rm=TRUE)
test$Bsmt.Half.Bath[is.na(test$Bsmt.Half.Bath)]<-median(test$Bsmt.Half.Bath,na.rm=TRUE)
test$BsmtFin.SF.1[is.na(test$BsmtFin.SF.1)]<-median(test$BsmtFin.SF.1,na.rm=TRUE)
test$BsmtFin.SF.2[is.na(test$BsmtFin.SF.2)]<-median(test$BsmtFin.SF.2,na.rm=TRUE)
test$Bsmt.Unf.SF[is.na(test$Bsmt.Unf.SF)]<-median(test$Bsmt.Unf.SF,na.rm=TRUE)
test$Total.Bsmt.SF[is.na(test$Total.Bsmt.SF)]<-median(test$Total.Bsmt.SF,na.rm=TRUE)
test$Garage.Cars[is.na(test$Garage.Cars)]<-median(test$Garage.Cars,na.rm=TRUE)
test$Garage.Area[is.na(test$Garage.Area)]<-median(test$Garage.Area,na.rm=TRUE)
```

No null values for train or test set
```{r}
sapply(train, function(y) sum(length(which(is.na(y)))))
```

```{r}
sapply(test, function(y) sum(length(which(is.na(y)))))
```

##### MS Subclass

Change type of MS Subclass to Factor
```{r}
train$MS.SubClass <- as.factor(train$MS.SubClass)
test$MS.SubClass <- as.factor(test$MS.SubClass)
```

All except RL, RM, and FV are infrequent for MS Zoning, so include these and an "other" category.
```{r}
table(train$MS.Zoning)
```

##### MS Zoning

```{r}
train$MS.Zoning <- ifelse(train$MS.Zoning %in% c("A (agr)", "C (all)", "I (all)", "RH", "RM"), "other", train$MS.Zoning)
test$MS.Zoning <- ifelse(test$MS.Zoning %in% c("A (agr)", "C (all)", "I (all)", "RH", "RM"), "other", test$MS.Zoning)
```

```{r}
table(train$MS.Zoning)
```

##### Alley

Dummy Variable for has.alley: seems unlikely type of alley matters, so drop original alley column
```{r}
train$Has.Alley <- as.factor(ifelse(train$Alley == "None", 0, 1))
test$Has.Alley <- as.factor(ifelse(test$Alley == "None", 0, 1))
train <- select(train, -Alley)
test <- select(test, -Alley)
```

```{r}
table(train$has.alley)
```

##### Lot.Shape

Lot Shape is ordinal, so we want to have values that are numeric to represent the ordering.
```{r}
train$Lot.Shape <- case_when(train$Lot.Shape == "Reg" ~ 0,
                             train$Lot.Shape == "IR1" ~ 1,
                             train$Lot.Shape == "IR2" ~ 2,
                             train$Lot.Shape == "IR3" ~ 3)
test$Lot.Shape <- case_when(test$Lot.Shape == "Reg" ~ 0,
                             test$Lot.Shape == "IR1" ~ 1,
                             test$Lot.Shape == "IR2" ~ 2,
                             test$Lot.Shape == "IR3" ~ 3)
```

#### Lot Config

```{r}
table(train$Lot.Config)
```

Dummy for Inside lot or not, as that is the most frequent value.
```{r}
train$Lot.Config <- ifelse(train$Lot.Config == "Inside", "Inside", "other")
test$Lot.Config <- ifelse(test$Lot.Config == "Inside", "Inside", "other")
```

```{r}
table(train$Lot.Config)
```

##### Land Slope

Again, Land Slope in ordinal, so encode appropriately.
```{r}
table(train$Land.Slope)
```

```{r}
train$Land.Slope <- case_when(train$Land.Slope == "Gtl" ~ 0,
                              train$Land.Slope == "Mod" ~ 1,
                              train$Land.Slope == "Sev" ~ 2)
test$Land.Slope <- case_when(test$Land.Slope == "Gtl" ~ 0,
                              test$Land.Slope == "Mod" ~ 1,
                              test$Land.Slope == "Sev" ~ 2)
```


##### Condition.1

Condition 1 has mostly normal values, so include a dummy for normal or not.
```{r}
table(train$Condition.1)
```

```{r}
train$Condition.1 <- ifelse(train$Condition.1 == "Norm", "norm", "other")
test$Condition.1 <- ifelse(test$Condition.1 == "Norm", "norm", "other")
```

```{r}
table(train$Condition.1)
```

##### Condition.2

Condition 2 has almost all normal values, so it will include only a dummy variable for normal or not.
```{r}
table(train$Condition.2)
```

```{r}
train$Condition.2 <- as.factor(ifelse(train$Condition.2 == "Norm", "norm", "other"))
test$Condition.2 <- as.factor(ifelse(test$Condition.2 == "Norm", "norm", "other"))
```

```{r}
table(train$Condition.2)
```

##### Roof Style

Only Gable and Hip have significant numbers, so we can include an "other" category for the rest.
```{r}
train$Roof.Style <- ifelse(train$Roof.Style %in% c("Flat", "Gambrel", "Mansard", "Shed"), "other", train$Roof.Style)
test$Roof.Style <- ifelse(test$Roof.Style %in% c("Flat", "Gambrel", "Mansard", "Shed"), "other", train$Roof.Style)
```

```{r}
table(train$Roof.Style)
```

###### Roof Matl

```{r}
table(train$Roof.Matl)
```

Include dummy variable for CompShg, as it is the only significant quantity
```{r}
train$Roof.Matl <- ifelse(train$Roof.Matl == "CompShg", "CompShg", "other")
test$Roof.Matl <- ifelse(test$Roof.Matl == "CompShg", "CompShg", "other")
```

```{r}
table(train$Roof.Matl)
```

##### Exterior 1

Mostly Vinyl, HdBoard, and Plywood, so include these and a dummy for others
```{r}
table(train$Exterior.1st)
```

```{r}
train$Exterior.1st <- ifelse(train$Exterior.1st %in% c("VinylSd", "HdBoard", "Plywood"), train$Exterior.1st, "other")
test$Exterior.1st <- ifelse(test$Exterior.1st %in% c("VinylSd", "HdBoard", "Plywood"), test$Exterior.1st, "other")
```

```{r}
table(train$Exterior.1st)
```

##### Exterior 2

Similar to exterior 1, but CmentBd, HdBoard, MetalSd, Plywood, VinylSd, and Wd Sdng are high in frequency.
```{r}
table(train$Exterior.2nd)
```

```{r}
train$Exterior.2nd <- ifelse(train$Exterior.2nd %in% c("CmentBd", "HdBoard", "MetalSd", "Plywood", "VinylSd", "Wd Sdng"), train$Exterior.2nd, "other")
test$Exterior.2nd <- ifelse(test$Exterior.2nd %in% c("CmentBd", "HdBoard", "MetalSd", "Plywood", "VinylSd", "Wd Sdng"), test$Exterior.2nd, "other")
```

```{r}
table(train$Exterior.2nd)
```

##### Masonry Vnr Type

The only significant values are BrkFace and None. Include other for the rest.
```{r}
table(train$Mas.Vnr.Type)
```

```{r}
train$Mas.Vnr.Type <- ifelse(train$Mas.Vnr.Type %in% c("BrkFace", "None"), train$Mas.Vnr.Type, "other")
test$Mas.Vnr.Type <- ifelse(test$Mas.Vnr.Type %in% c("BrkFace", "None"), test$Mas.Vnr.Type, "other")
```

```{r}
table(train$Mas.Vnr.Type)
```

##### Exter Qual

```{r}
table(train$Exter.Qual)
```


Again, ordinal, so encode appropriately.
```{r}
train$Exter.Qual <- case_when(train$Exter.Qual == "Ex" ~ 4,
                              train$Exter.Qual == "Gd" ~ 3,
                              train$Exter.Qual == "TA" ~ 2,
                              train$Exter.Qual == "Fa" ~ 1,
                              train$Exter.Qual == "Po" ~ 0)
test$Exter.Qual <- case_when(test$Exter.Qual == "Ex" ~ 4,
                              test$Exter.Qual == "Gd" ~ 3,
                              test$Exter.Qual == "TA" ~ 2,
                              test$Exter.Qual == "Fa" ~ 1,
                              test$Exter.Qual == "Po" ~ 0)
```

```{r}
table(train$Exter.Qual)
```

##### Foundation

Slab, Stone, and Wood are infrequent, so replace those with other.
```{r}
table(train$Foundation)
```

```{r}
train$Foundation <- ifelse(train$Foundation %in% c("BrkTil", "CBlock", "PConc"), train$Foundation, "other")
test$Foundation <- ifelse(test$Foundation %in% c("BrkTil", "CBlock", "PConc"), test$Foundation, "other")
```

```{r}
table(train$Foundation)
```

##### Create Dummy Variable for whether house has a basement

```{r}
train$Has.Basement <- ifelse(train$Bsmt.Qual == "No Basement", 0, 1)
test$Has.Basement <- ifelse(test$Bsmt.Qual == "No Basement", 0, 1)
```


##### Bsmt Qual

```{r}
table(train$Bsmt.Qual)
```


Again, Ordinal. This time, not having a basement is zero, as that seems to make sense in the heirarchy.
```{r}
train$Bsmt.Qual <- case_when(train$Bsmt.Qual == "Ex" ~ 5,
                             train$Bsmt.Qual == "Gd" ~ 4,
                             train$Bsmt.Qual == "TA" ~ 3,
                             train$Bsmt.Qual == "Fa" ~ 2,
                             train$Bsmt.Qual == "Po" ~ 1,
                             train$Bsmt.Qual == "No Basement" ~ 0)
test$Bsmt.Qual <- case_when(test$Bsmt.Qual == "Ex" ~ 5,
                             test$Bsmt.Qual == "Gd" ~ 4,
                             test$Bsmt.Qual == "TA" ~ 3,
                             test$Bsmt.Qual == "Fa" ~ 2,
                             test$Bsmt.Qual == "Po" ~ 1,
                             test$Bsmt.Qual == "No Basement" ~ 0)
```

```{r}
table(train$Bsmt.Qual)
```


##### Bsmt Cond

Again, ordinal, and represent no basement as zero.
```{r}
table(train$Bsmt.Cond)
```

```{r}
train$Bsmt.Cond <- case_when(train$Bsmt.Cond == "Ex" ~ 5,
                             train$Bsmt.Cond == "Gd" ~ 4,
                             train$Bsmt.Cond == "TA" ~ 3,
                             train$Bsmt.Cond == "Fa" ~ 2,
                             train$Bsmt.Cond == "Po" ~ 1,
                             train$Bsmt.Cond == "No Basement" ~ 0)
test$Bsmt.Cond <- case_when(test$Bsmt.Cond == "Ex" ~ 5,
                             test$Bsmt.Cond == "Gd" ~ 4,
                             test$Bsmt.Cond == "TA" ~ 3,
                             test$Bsmt.Cond == "Fa" ~ 2,
                             test$Bsmt.Cond == "Po" ~ 1,
                             test$Bsmt.Cond == "No Basement" ~ 0)
```

```{r}
table(train$Bsmt.Cond)
```

##### Bsmt Exposure

Again, ordinal with No Basement as 0.
```{r}
table(train$Bsmt.Exposure)
```

```{r}
train$Bsmt.Exposure <- case_when(train$Bsmt.Exposure == "Gd" ~ 4,
                                 train$Bsmt.Exposure == "Av" ~ 3,
                                 train$Bsmt.Exposure == "Mn" ~ 2,
                                 train$Bsmt.Exposure == "No" ~ 1,
                                 train$Bsmt.Exposure == "No Basement" ~ 0)
test$Bsmt.Exposure <- case_when(test$Bsmt.Exposure == "Gd" ~ 4,
                                 test$Bsmt.Exposure == "Av" ~ 3,
                                 test$Bsmt.Exposure == "Mn" ~ 2,
                                 test$Bsmt.Exposure == "No" ~ 1,
                                 test$Bsmt.Exposure == "No Basement" ~ 0)
```

```{r}
table(train$Bsmt.Exposure)
```

##### BsmtFin Type 2

Although Type 1 had an even distribution, type 2 has largely unfinished.
```{r}
table(train$BsmtFin.Type.2)
```

```{r}
train$BsmtFin.Type.2 <- ifelse(train$BsmtFin.Type.2 == "Unf", "Unf", "other")
test$BsmtFin.Type.2 <- ifelse(test$BsmtFin.Type.2 == "Unf", "Unf", "other")
```

```{r}
table(train$BsmtFin.Type.2)
```

##### Heating

Dummy variable for GasA
```{r}
table(train$Heating)
```

```{r}
train$Heating <- ifelse(train$Heating == "GasA", "GasA", "other")
test$Heating <- ifelse(test$Heating == "GasA", "GasA", "other")
```

```{r}
table(train$Heating)
```

##### Heating QC

Ordinal, so encode appropriately.
```{r}
table(train$Heating.QC)
```


```{r}
train$Heating.QC <- case_when(train$Heating.QC == "Ex" ~ 4,
                              train$Heating.QC == "Gd" ~ 3,
                              train$Heating.QC == "TA" ~ 2,
                              train$Heating.QC == "Fa" ~ 1,
                              train$Heating.QC == "Po" ~ 0)
test$Heating.QC <- case_when(test$Heating.QC == "Ex" ~ 4,
                              test$Heating.QC == "Gd" ~ 3,
                              test$Heating.QC == "TA" ~ 2,
                              test$Heating.QC == "Fa" ~ 1,
                              test$Heating.QC == "Po" ~ 0)
```

```{r}
table(train$Heating.QC)
```

##### Electrical

Only include Sbrkr or other.
```{r}
table(train$Electrical)
```

```{r}
train$Electrical <- ifelse(train$Electrical == "SBrkr", "SBrkr", "other")
test$Electrical <- ifelse(test$Electrical == "SBrkr", "SBrkr", "other")
```

```{r}
table(train$Electrical)
```

```{r}
train$Kitchen.Qual <- case_when(train$Kitchen.Qual == "Ex" ~ 4,
                                train$Kitchen.Qual == "Gd" ~ 3,
                                train$Kitchen.Qual == "TA" ~ 2,
                                train$Kitchen.Qual == "Fa" ~ 1,
                                train$Kitchen.Qual == "Po" ~ 0)
test$Kitchen.Qual <- case_when(test$Kitchen.Qual == "Ex" ~ 4,
                                test$Kitchen.Qual == "Gd" ~ 3,
                                test$Kitchen.Qual == "TA" ~ 2,
                                test$Kitchen.Qual == "Fa" ~ 1,
                                test$Kitchen.Qual == "Po" ~ 0)
```

```{r}
table(train$Kitchen.Qual)
```

##### Functional

Dummy variable for functionality less than typical
```{r}
table(train$Functional)
```

```{r}
train$Functional <- ifelse(train$Functional == "Typ", "Typ", "BelowTyp")
test$Functional <- ifelse(test$Functional == "Typ", "Typ", "BelowTyp")
```

```{r}
table(train$Functional)
```

##### Create Dummy Variable for has fireplace

```{r}
train$Has.Fireplace <- ifelse(train$Fireplace.Qu == "None", 1, 0)
test$Has.Fireplace <- ifelse(test$Fireplace.Qu == "None", 1, 0)
```

##### Fireplace Qu

```{r}
train$Fireplace.Qu <- case_when(train$Fireplace.Qu == "Ex" ~ 5,
                                train$Fireplace.Qu == "Gd" ~ 4,
                                train$Fireplace.Qu == "TA" ~ 3,
                                train$Fireplace.Qu == "Fa" ~ 2,
                                train$Fireplace.Qu == "Po" ~ 1,
                                train$Fireplace.Qu == "None" ~ 0)
test$Fireplace.Qu <- case_when(test$Fireplace.Qu == "Ex" ~ 5,
                                test$Fireplace.Qu == "Gd" ~ 4,
                                test$Fireplace.Qu == "TA" ~ 3,
                                test$Fireplace.Qu == "Fa" ~ 2,
                                test$Fireplace.Qu == "Po" ~ 1,
                                test$Fireplace.Qu == "None" ~ 0)
```

```{r}
table(train$Fireplace.Qu)
```

##### Dummy Variable for has garage

```{r}
train$Has.Garage <- ifelse(train$Garage.Type == "No Garage", 1, 0)
test$Has.Garage <- ifelse(test$Garage.Type == "No Garage", 1, 0)
```

##### Garage Finish

Ordinal, so encode accordingly.
```{r}
table(train$Garage.Finish)
```

```{r}
train$Garage.Finish <- case_when(train$Garage.Finish == "Fin" ~ 3,
                                 train$Garage.Finish == "RFn" ~ 2,
                                 train$Garage.Finish == "Unf" ~ 1,
                                 train$Garage.Finish == "No Garage" ~ 0)
test$Garage.Finish <- case_when(test$Garage.Finish == "Fin" ~ 3,
                                 test$Garage.Finish == "RFn" ~ 2,
                                 test$Garage.Finish == "Unf" ~ 1,
                                 test$Garage.Finish == "No Garage" ~ 0)
```

```{r}
table(train$Garage.Finish)
```

##### Garage Quality

```{r}
table(train$Garage.Qual)
```

```{r}
train$Garage.Qual <- case_when(train$Garage.Qual == "Ex" ~ 5,
                               train$Garage.Qual == "Gd" ~ 4,
                               train$Garage.Qual == "TA" ~ 3,
                               train$Garage.Qual == "Fa" ~ 2,
                               train$Garage.Qual == "Po" ~ 1,
                               train$Garage.Qual == "No Garage" ~ 0)
test$Garage.Qual <- case_when(test$Garage.Qual == "Ex" ~ 5,
                               test$Garage.Qual == "Gd" ~ 4,
                               test$Garage.Qual == "TA" ~ 3,
                               test$Garage.Qual == "Fa" ~ 2,
                               test$Garage.Qual == "Po" ~ 1,
                               test$Garage.Qual == "No Garage" ~ 0)
```

```{r}
table(train$Garage.Qual)
```

#### Paved Drive

There are not many unpaved, so only include a dummy for paved or unpaved.
```{r}
table(train$Paved.Drive)
```

```{r}
train$Paved.Drive <- ifelse(train$Paved.Drive == "Y", "Paved", "Unpaved")
test$Paved.Drive <- ifelse(test$Paved.Drive == "Y", "Paved", "Unpaved")
```

```{r}
table(train$Paved.Drive)
```

##### Pool Area and Quality don't appear relevant because most don't have a pool; create a dummy for having a pool instead

```{r}
train$Has.Pool <- ifelse(train$Pool.QC == "No Pool", 0, 1)
test$Has.Pool <- ifelse(test$Pool.QC == "No Pool", 0, 1)
```


Drop columns:
```{r}
train <- select(train, -Pool.QC, -Pool.Area)
test <- select(test, -Pool.QC, -Pool.Area)
```

##### Fence

Ordinal, so encode appropriately with no fence as 0.
```{r}
table(train$Fence)
```


```{r}
train$Fence <- case_when(train$Fence == "GdPrv" ~ 4,
                         train$Fence == "McPrv" ~ 3,
                         train$Fence == "GdWo" ~ 2,
                         train$Fence == "McWw" ~ 1,
                         train$Fence == "No Fence" ~ 0)
test$Fence <- case_when(test$Fence == "GdPrv" ~ 4,
                         test$Fence == "McPrv" ~ 3,
                         test$Fence == "GdWo" ~ 2,
                         test$Fence == "McWw" ~ 1,
                         test$Fence == "No Fence" ~ 0)
```

```{r}
table(train$Fence)
```


##### Misc Feature

Create dummy for having a miscellaneous feature
```{r}
table(train$Misc.Feature)
```

```{r}
train$Misc.Feature <- ifelse(train$Misc.Feature == "None", 0, 1)
test$Misc.Feature <- ifelse(test$Misc.Feature == "None", 0, 1)
```

```{r}
table(train$Misc.Feature)
```

##### Sale Type

Eliminate sparse categories, those other than New and Wd
```{r}
table(train$Sale.Type)
```

```{r}
train$Sale.Type <- ifelse(train$Sale.Type %in% c("New", "WD "), train$Sale.Type, "other")
test$Sale.Type <- ifelse(test$Sale.Type %in% c("New", "WD "), test$Sale.Type, "other")
```

```{r}
table(train$Sale.Type)
```

##### Filter > 4000 sq. ft. observations

```{r}
train <- train[train$Gr.Liv.Area < 4000,]
```

##### Eliminate Irrelevant Variables

Utilities
```{r}
table(train$Utilities)
```

```{r}
train <- train %>% select(-Utilities)
test <- test %>% select(-Utilities)
```

Wood Deck Sq Ft.: many zero values, so include a dummy instead

```{r}
length(train$Wood.Deck.SF[train$Wood.Deck.SF == 0])
```

```{r}
train$Has.Wood.Deck <- ifelse(train$Wood.Deck.SF == 0, 0, 1)
test$Has.Wood.Deck <- ifelse(test$Wood.Deck.SF == 0, 0, 1)
```

Open Porch Sq. Ft.: Same

```{r}
length(train$Open.Porch.SF[train$Open.Porch.SF == 0])
```

```{r}
train$Has.Open.Porch <- ifelse(train$Open.Porch.SF == 0, 0, 1)
test$Has.Open.Porch <- ifelse(test$Open.Porch.SF == 0, 0, 1)
```

Enclosed Porch

```{r}
length(train$Enclosed.Porch[train$Enclosed.Porch == 0])
```

```{r}
train$Has.Enclosed.Porch <- ifelse(train$Enclosed.Porch == 0, 0, 1)
test$Has.Enclosed.Porch <- ifelse(test$Enclosed.Porch == 0, 0, 1)
```

3-Ssn Porch
```{r}
length(train$X3Ssn.Porch[train$X3Ssn.Porch == 0])
```

```{r}
train$Has.X3Ssn.Porch <- ifelse(train$X3Ssn.Porch == 0 , 0, 1)
test$Has.X3Ssn.Porch <- ifelse(test$X3Ssn.Porch == 0 , 0, 1)
```

Screen Porch
```{r}
length(train$Screen.Porch[train$Screen.Porch == 0])
```

```{r}
train$Has.Screen.Porch <- ifelse(train$Screen.Porch == 0, 0, 1)
test$Has.Screen.Porch <- ifelse(test$Screen.Porch == 0, 0, 1)
```

Remove Variables:

```{r}
train <- train %>% select(-Wood.Deck.SF, -Open.Porch.SF, -Enclosed.Porch, -X3Ssn.Porch, -Screen.Porch)
test <- test %>% select(-Wood.Deck.SF, -Open.Porch.SF, -Enclosed.Porch, -X3Ssn.Porch, -Screen.Porch)
```





